---
title: 'R4DataScience Repository'
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  fig.path = "figures/README-"
)
```

## Overview

This repository includes a bunch of amound of codes learned from:

* <https://r4ds.had.co.nz/data-visualisation.html>
* <https://www.amazon.co.jp/dp/1461471370/ref=cm_sw_em_r_mt_dp_U_UeVmCbNT4X2J0>
* <https://tensorflow.rstudio.com/>

Some Useful Dataset Related Resources:

* <http://www.sthda.com/english/wiki/r-built-in-data-sets>

## Environment Installation

About [Packrat](https://rstudio.github.io/packrat/), the walkthrough document seems the best document to follow after cloning: <https://rstudio.github.io/packrat/walkthrough.html>

Although I have set `Packrat` automatically snapshot changes, it seems that changes not updated into `packrat.lock` file as expectation.

If you load this repository as a R project into your R studio but not see packrat automatically restoring package dependencies, follow the commands below:

```{r eval=FALSE}
packrat::restore()
```

Once new packages installed,
```{r eval=FALSE}
packrat::snapshot()
```

Check and cleaning up actually unused package at the commit or push time,
```{r eval=FALSE}
packrat::status()

packrat::clean()
```

