---
title: "Eruptions Data of Old Faithful Geyser"
author: "Richard Lea <https://github.com/chigix>"
date: "2019/1/3"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Dataset

```{r dataset}
faithful
```

## Observation on the `eruption`

```{r eruption-observation-count}
faithful %>%
  mutate(
    erupt_group = cut_width(eruptions, width = 0.25)
  ) %>%
  ggplot() + 
  geom_histogram(mapping = aes(x = eruptions, fill = erupt_group), binwidth = 0.25)
```

## Observation on the `waiting time`

```{r}
faithful %>%
  ggplot() + 
  geom_freqpoly(mapping = aes(x = waiting, color = cut_width(eruptions, width = 0.25)), binwidth = 10)
```
This graph above shows there are more observations when waiting arount 50~60 mins and 77~82mins than other length of waiting.

The graph above is exploring the distribution of `waiting`, a continuous variable, broken down by the categorical variable, `erupt_group`.

## Examine Relationship

```{r}
ggplot(faithful, mapping = aes(x = waiting, y = eruptions)) +
  geom_point()
```
The scatterplot above shows the pattern, as well, that longer wait times are associated with longer eruptions, the clues about relationships are called **Patterns**.

Two **Clusters** are displayed through the scatterplot too.

The Patterns are extracted by modeling.
