---
title: "ggplot2::diamonds Data Case"
author: "Richard Lea"
date: "2018/12/16"
bibliography: bibliography.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
require(ggplot2)
require(dplyr)
require(scales)
```

## DataSet

```{r diamonds}
ggplot2::diamonds
```

## Basic Statistical Transformations: `Stat`

```{r stat_summary}
ggplot(data = diamonds) +
  stat_summary(
    mapping = aes(x = cut, y = depth),
    fun.ymin = min,
    fun.ymax = max,
    fun.y = median
  )
```
```{r identity_position_in_bar_graph}
# Max value is less than default stack position
ggplot(diamonds, mapping = aes(x = cut)) +
  # geom_bar(mapping = aes(fill = clarity), alpha = 2/5, position = "identity")
  geom_bar(mapping = aes(color = clarity), fill = NA, position = "identity")
```
```{r fill_position_in_bar_chart}
base_plot <- ggplot(diamonds, mapping = aes(x = cut))
# Easy for proportions comparing
base_plot + geom_bar(mapping = aes(fill = clarity), position = "fill")
# Separate overlapping objects
base_plot + geom_bar(mapping = aes(fill = clarity), position = "dodge")
```
```{r jitter_for_overplotting_in_scatterplot}
ggplot(diamonds) +
  # geom_point(mapping = aes(x = depth, y = price))
                                    # Many points around depth of 60 overlapped
  # geom_point(data = filter(diamonds, depth == 60), mapping = aes(x = depth, y = price))
                                    # Still hard to see where the mass of data is
  # geom_point(data = filter(diamonds, depth == 60), mapping = aes(x = depth, y = price), position = "jitter")
                                    # Spreads 511 points out making points receive random noises
  geom_jitter(data = filter(diamonds, depth == 60), mapping = aes(x = depth, y = price), width = 0.1)
                                    # Make the total spread to be 0.2 around 60
cat("Filtered Result Rows:", NROW(filter(diamonds, depth == 60)))
diamonds %>% filter(depth == 60) %>% count(cut)
```

### `cut`: Examine the distribution of a categorical variable

A variable is **categorical** if it can only take one of a small set of values.[@wickham2010]

```{r basic-count}
(
  ggplot(data = diamonds) +
    geom_bar(mapping = aes(x = cut, y = ..count..))
    # geom_bar(mapping = aes(x = cut))  # Implicit Count Computing
    # Error: binwidth is not supported in bar_graph ↓
    # geom_bar(mapping = aes(x = cut), binwidth = 30)
    # stat_count(mapping = aes(x = cut))       # count transformation underlying
)
```

What's more, preset factors on the variables in the tibble make possible values of cuts be shown in order from *Fair* through *Ideal* in the graph above, which could available through `count` like the following table as well:

```{r direct-count-on-cut}
diamonds %>% count(cut)
```
```{r proportion-computing}
(
  ggplot(diamonds) +
    # geom_bar(mapping = aes(x = cut, y = ..prop..))
    # Groupwise Proportion will make it work on default grouping strategy
    # which is grouped by color variable, in this data case
    geom_bar(mapping = aes(x = cut, y = ..prop.., fill = color))
    # Test if the color variable used as group definition
)
(
  ggplot(diamonds) +
    # disable all grouping strategy, makes ggplot take
    # the whole dataset as one group
    geom_bar(mapping = aes(x = cut, y = ..prop.., group = 1))
)
```
```{r stacked-bar-graph}
(
  ggplot(diamonds) +
    # geom_bar(mapping = aes(x = cut, y = ..count.., fill = clarity), position = "stack")
    geom_bar(mapping = aes(x = cut, y = ..count.., fill = clarity))
    # stack is default position in geom_bar function
)
(
  ggplot(diamonds) +
    geom_bar(mapping = aes(x = cut, y = ..prop.., fill = clarity, group = 1), position = "stack")
    # position does not work in proportion plot
)
```
```{r stat_count_in_bar_graph}
ggplot(diamonds) +
  geom_bar(mapping = aes(x = cut, color = cut)) # Color a bar chart
ggplot(diamonds) +
  geom_bar(mapping = aes(x = cut, fill = cut))  # Fill a bar chart
```
```{r stat_identity_in_bar_graph}
(demo_small_table <- tibble::tribble(
  ~cut,         ~freq,
  #-----------|------
  "Fair",       1000,      # Key in a table should be unique for `identity stat`
  "Good",       2000,
  "Very Good",  3000,
  "Premium",    4000,
  "Ideal",      5000
))
demo_small_table %>%
  mutate(cut = fct_inorder(cut)) %>%
  rename(
    `カット` = cut,
    `頻度` = freq
  ) %>%
  ggplot() +
    geom_bar(mapping = aes(x = `カット`, y = `頻度`), stat = "identity")
```

### `price`: Examine the distribution of a continuous variable

Although Bar-Chart is available for continuous variable, but `stat_bin` is not supported by `geom_bar`, which makes `geom_histogram` or `geom_freqpoly` functions is much more used for continuous variable, exploring a variety of binwidths.

```{r stat_bin}
ggplot(data = diamonds)+
  geom_histogram(mapping = aes(price, fill = cut), binwidth = 500)
# Get the tabular form data for continuous data
count(diamonds, cut_width(price, width = 500))
```
In the graph above, the tallest bar shows that almost 10,000 observations have a `price` value between 750 and 1750(the range is much more easy to see in the tabular summary), where

* `r 750` is called the left edge
* `r 1750` is called the right edge

of that bar.

### `cut`~`price`: Covariation between a categorical variable and continuous variable

```{r group-continuous-categorical-var-with-bars}
ggplot(diamonds) +
  geom_histogram(aes(x = price, fill = cut), bins = 30)
```
```{r group-continuous-categorical-var-with-lines}
ggplot(diamonds) +
  geom_freqpoly(aes(x = price, color = cut), bins = 30)
```

Instead of displaying count, plotting in density is an approach as well, which make sure that the area under each frequency polygon is one.
```{r show-in-density}
ggplot(diamonds) +
  geom_freqpoly(aes(x = price, y = ..density.., color = cut), bins = 30)
```

### `cut`~`color`: Covariantion between Two categorical Variables

```{r geom_count}
ggplot(diamonds) +
  geom_count(mapping = aes(x = cut, y = color))
## This computing could also be available in table form:
diamonds %>% count(color, cut)
```
```{r geom_tile}
diamonds %>%
  count(color, cut) %>%
  ggplot() +
  geom_tile(mapping = aes(x = cut, y = color, fill = n)) +
  scale_fill_continuous(trans = 'reverse') +
  guides(fill = guide_legend(reverse = T))
```

### `carat`~`price`: Covariation between Two Continuous Variables

**Scatterplots** become less useful as the size of the dataset grows, because of overplotting among points. Event **transparency** is still challenging for very large datasets.

```{r transforming_on_scatterplot_of_price_vs_carat}
(price_vs_carat <- ggplot(data = diamonds, mapping = aes(x = carat, y = price)) +
  geom_point(alpha = 1/10))
price_vs_carat + 
  scale_x_log10(
    breaks = trans_breaks("log10", function(x)  10 ^ x),
    labels = trans_format("log10", math_format(10 ^ .x))
  ) +
  scale_y_log10(
    breaks = trans_breaks("log10", function(x)  10 ^ x),
    labels = trans_format("log10", math_format(10 ^ .x))
  ) + geom_smooth(method = "lm")
price_vs_carat + coord_trans(x = "log10", y = "log10")
price_vs_carat + scale_x_log10() + scale_y_log10() + geom_smooth(method = "lm") + coord_polar()
```
```{r bin2d-visualization}
ggplot(diamonds) +
  geom_bin2d(mapping = aes(x = carat, y = price))
```
```{r hexbin-visualization}
ggplot(diamonds) +
  geom_hex(aes(x = carat, y = price))
```
```{r bin-one-var-to-categorical}
ggplot(diamonds) +
  geom_boxplot(mapping = aes(x = carat, y = price, group = cut_width(carat, 0.5)))
```
```{r display-same-number-points-in-each-bin}
ggplot(diamonds) +
  geom_boxplot(mapping = aes(
    x = carat,
    y = price,
    group = cut_number(carat, 20)
  ))
```
```{r}
model <- lm(log(price) ~ log(carat), data = diamonds)
diamonds %>%
  add_residuals(model = model) %>%
  mutate(resid = exp(resid)) %>%
  ggplot() + geom_point(mapping = aes(x = carat, y = resid))
```
```{r}
model <- lm(log(price) ~ log(carat), data = diamonds)
diamonds %>%
  add_residuals(model = model) %>%
  mutate(resid = exp(resid)) %>%
  ggplot() + geom_boxplot(mapping = aes(x = cut, y = resid))
```
Residual = Observed y-value - Predicted y-value


## Coordinate Systems

### Coordinate Transformation

A regular bar chart converted into polar coordinates produces another type of graphic: the Coxcomb plot [@wickham2010].
```{r coxcomb_plot_on_cut_variable}
bar <- ggplot(diamonds) +
  geom_bar(
    mapping = aes(x = cut, fill = cut),
    show.legend = FALSE,     # Because x-axis has represented meanings of each bar
    width = 1
  )
bar <- bar + theme(aspect.ratio = 1)   # Set aspect ratio of the panel to the picture
(bar <- bar + labs(x = NULL, y = NULL))
bar + coord_flip()
bar + coord_polar()
```
```{r proportion_one_variable}
(bar_plot <-
   ggplot(diamonds, mapping = aes(x = "", fill = clarity)) +
   geom_bar(width = 1))
bar_plot + coord_polar(theta = "y")
bar_plot + coord_polar(theta = "x")
```

### Zooming

```{r}
(beforeZoom <- ggplot(diamonds) +
  geom_histogram(mapping = aes(x = y), binwidth = 0.5))
beforeZoom + coord_cartesian(ylim = c(0, 30))
beforeZoom + coord_cartesian(xlim = c(-2, 4), ylim = c(0, 10))
```

Through the graph above, filter records by the variable of y and see them in a detailed table:

```{r}
diamonds %>%
  filter(y < 3 | y > 20) %>%
  select(price, x, y, z) %>%
  arrange(y)
```

For the values that are unusual, replaceing them with missing values is a recommended practice, after which the graph created again could be less affected by outliers.

```{r}
diamonds %>%
  mutate(
    y = ifelse(y < 3 | y > 20, NA, y) # Remove the effect from unusual values
  ) %>%
  ggplot() +
  geom_histogram(mapping = aes(x = y), binwidth = 0.5)
```

## References
