---
title: "nlme::Oxboys data case"
author: "Richard Lea <https://github.com/chigix>"
date: "2019/1/3"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library("nlme")
```

## Dataset

This dataset records the heights and centered ages of 26 boys, measured on nine occasions, which result in the data frame consists $26\ \text{Subject} \times 9\ \text{Occasion} = 234$ rows.

```{r dataset}
as_tibble(nlme::Oxboys)
```

## Grouping Visualization

```{r single-line-connect-all-observations-problem}
ggplot(nlme::Oxboys, aes(age, height)) +
  geom_line()
```

This Line Plot above looks weird because all of the observations are treated as one continuous data set, rather than grouping them into 26 subjects(actually 26 boys).

```{r fix-line-plot-with-subjects-grouping}
ggplot(Oxboys, aes(age, height)) +
  geom_line(mapping = aes(group = Subject))
```
Now, 26 lines of data corresponding to 26 subjects(actually 26 boys) are obtained by defining the grouping structure explicitly.

```{r grouping-in-geom-smooth-problem}
ggplot(Oxboys, aes(age, height)) +
  geom_line(mapping = aes(group = Subject)) +
  # Use Linear Models
  geom_smooth(mapping = aes(group = Subject), method = "lm", se = FALSE)
```
```{r change-group-aesthetic-to-fit-across-all-boys}
ggplot(Oxboys, aes(age, height)) +
  geom_line(mapping = aes(group = Subject)) +
  # Use Linear Models
  # geom_smooth(mapping = aes(), size = 2, method = "lm", se = FALSE)
  # group = 1 disable all grouping strategy, which makes smooth function here
  #           to try to create a conditional means connecting all occasions
  geom_smooth(mapping = aes(group = 1), size = 2, method = "lm", se = FALSE)
```
```{r boxplot-visualize-each-occasion}
ggplot(Oxboys, aes(Occasion, height)) +
  geom_boxplot() +
  geom_point(aes(color = Subject))
```
The result shows that `group` aesthetic is not needed to specify, because the default grouping works in `boxplot` for the discrete occasion variable.
