---
title: "New York City Flights Data Playground"
---

```{r setup, include=FALSE}
library("nycflights13")
library("tidyverse")
library("lubridate")
library("modelr")

options(na.action = na.warn)
```

## Dataset

```{r arrange_rows}
flights
arrange(flights, year, desc(month), day)
```
Take a simple summary of this dataset:
```{r summary_for_nycflights_dataset}
str(flights)
```

### Common Variable Types

Common Variable Types in DataFrame:

* `int` stands for integers
* `dbl` stands for doubles or real numbers
* `chr` stands for character vectors, where strings are included as well
* `dttm` is the abbreviation for `data-times`
* `date` stands for dates
* `lgl` stands for logical or vectors that contain `TRUE` or `FALSE`
* `fctr` stands for factors, which R uses to represent categorical variables with fixed possible values

## `dplyr` Functions

```{r filter_rows}
# Select all flights on Jan 1st.
filter(flights, month == 1, day == 1)
# Select all flights in Nov or Dec.
filter(flights, month == 11 | month == 12) # Pay attention to the `|` here, it is not `||`
filter(flights, month %in% c(9, 10))        # Select all flights in Sep or Oct.
filter(flights, between(dep_time, 0, 600))  # Departed between midnight and 6am
```
```{r near_usage_cases}
cat("[ sqrt(2) ^ 2 == 2 ] ->", sqrt(2) ^ 2 == 2)
cat("\n[ near(sqrt(2) ^ 2, 2) ] ->", near(sqrt(2) ^ 2, 2))
cat("\n[ 1/49 * 49 == 1 ] ->", 1/49 * 49 == 1)
cat("\n[ near(1/49 * 49, 1) ] ->", near(1/49 * 49, 1))
```
```{r De_Morgans_raw}
# Pay attention to the `|` here, which isn't `||`
all(
  filter(flights, !(arr_delay > 120 | dep_delay > 120)) ==
    filter(flights, arr_delay <= 120, dep_delay <= 120)
)
```
```{r missing_values}
NA
NA > 33
NA - 2
NA ^ NA
is.na(NA)    # A recommended way to determine whether a value is missing
```
`NA` will make almost any operation unknown.
```{r select_columns}
select(flights, year:day)        # Same with the below
# select(flights, year, month, day)
select(flights, -(year: day))    # Select all colums except year, month and day
select(flights, starts_with("dep_"))
select(flights, contains("dep_"))
select(flights, contains("dep_"), everything()) # All dep fields moved to the start of the dataframe
select(flights, 出発時間 = dep_time, 予定出発 = sched_dep_time, 出発遅延 = dep_delay)
# Slight Difference between `select` and `rename`
rename(flights, 出発時間 = dep_time, 予定出発 = sched_dep_time, 出発遅延 = dep_delay)
```
```{r mutate_new_var}
mutate(
  select(flights, year:day, ends_with("delay"), distance, air_time),
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60,
  gain_per_hour = gain / air_time * 60 # Refer to the newly created column
)
transmute(
  flights,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60,
  gain_per_hour = gain / air_time * 60 # Refer to the newly created column
)   # which will drop all original variables in the table
```
```{r Grouped_summaries}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE), count = n())
summarise(
  group_by(flights, year, month, day),
  delay = mean(dep_delay, na.rm = TRUE),
  count = n()
)
```
```{r transformation_with_pipe}
# (delay_dataset <- summarise(
#   group_by(flights, dest),
#   count = n(),
#   dist = mean(distance, na.rm = TRUE),
#   delay = mean(arr_delay, na.rm = TRUE)
# ))
# (delay_filtered <- filter(delay_dataset, count > 20, dest != "HNL"))
#
# Use Pipe to make the transformation above concise:
(delays <- flights %>%
   group_by(dest) %>%
   summarise(
     count = n(),
     dist = mean(distance, na.rm = TRUE),
     delay = mean(arr_delay, na.rm = TRUE)
   ) %>%
   filter(count > 20, dest != "HNL")
)
ggplot(delays, aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```
```{r count_to_visualize_density_on_frequency_plot}
# Visualize the density through count
(delays <- flights
  %>% filter(!is.na(dep_delay),!is.na(arr_delay)) # Filter the flights not cancelled
  %>% group_by(tailnum)
  %>% summarise(
    delay = mean(arr_delay),
    count = n()
  )
)
filter(delays, count > 25)
delays %>%
  ggplot(mapping = aes(x = delay)) +
  geom_freqpoly(binwidth = 10) +
  xlab("delay [min]")
delays %>%
  # filter(count > 25) %>%     # Filter to few flights to see variation
  ggplot(mapping = aes(x = delay, y = count)) +
  geom_point(alpha = 1/8) +    # Use the stastistic through count
  xlab("delay [min]")
```
```{r arithmetic-mean-of-arrival-delays}
# Estimation Upon Average Delay among NYC-Flights
(
  flights
    %>% filter(!is.na(dep_delay),!is.na(arr_delay)) # Filter the flights not cancelled
    %>% group_by(year, month, day)
    %>% summarise(
      avg_delay_for_all = mean(arr_delay),
      avg_delay_for_positive_part = mean(arr_delay[arr_delay > 0])
    )
    %>% mutate(
      # date_field = make_datetime(year, month, day)
      date_field = ISOdate(year, month, day) # Attach date type variable
    )
    %>% ggplot()
          + geom_point(
            mapping = aes(x = date_field, y = avg_delay_for_all, colour = "DelayForAll")
          )
          + geom_point(
            mapping = aes(x = date_field, y = avg_delay_for_positive_part, colour = "DelayForPositive")
          )
          + scale_color_manual(
            # Set the order in the legend
            breaks = c("DelayForPositive", "DelayForAll"),
            values = c("DelayForAll" = "brown", "DelayForPositive" = "violetred1")
          )
)
```
```{r spread_of_distances}
(
  flights
    %>% filter(!is.na(dep_delay),!is.na(arr_delay)) # Filter the flights not cancelled
    %>% group_by(dest)
    %>% summarise(
      distanceAvg = mean(distance),
      min = min(distance),
      max = max(distance),
      median = median(distance),
      standardDeviation = sd(distance),
      interquartileRange = IQR(distance),
      interquartileRange_0 = quantile(distance, 0.75) - quantile(distance, 0.25),
      medianAbsoluteDeviation = mad(distance),
      prove_quantile1 = mean(distance < quantile(distance, 0.25)) # Get a Proportion
    )
    %>% arrange(desc(standardDeviation))
)
```

Measures of spread:

* Mean Squared Deviation: $\frac{\sum (x-\mu)^2}{N}$
    * $N$ = Number of Data
* Standard Deviation(Root Mean Squared Deviation):
  $\sigma = \sqrt{\frac{\sum (x-\mu)^2}{N}}$ 
    * $N$ = Number of Data
* Median Absolute Deviation:

    $MAD(D) = median({|d_i - median(D)|})$

```{r summary-distinct-values}
(
  flights
  %>% filter(!is.na(dep_delay),!is.na(arr_delay)) # Filter the flights not cancelled
  %>% group_by(dest)
  %>% summarise(
    count = n(),
    carriers = n_distinct(carrier), # Same as the below
    # carriers_0 = length(unique(carrier)),
    carriers_list = paste(unique(carrier), collapse = ", ")
  )
  %>% arrange(desc(carriers))
)
```

## Summary Fist and Last Flights each day

### Departure time only of first and last flights each day

```{r find-first-and-last-depart-time}
(
  flights
  %>% filter(!is.na(dep_delay),!is.na(arr_delay)) # Filter the flights not cancelled
  %>% group_by(year, month, day)
  %>% summarise(
    first_dep_a = min(dep_time),
    first_dep_b = first(dep_time),
    last_dep_a = max(dep_time),
    last_dep_b = last(dep_time),
    n_early = sum(dep_time < 500)   # Count flights left before 5am
  )
)
```

### Details of first and last flights each day

```{r ranking-summaries}
(
  flights
  %>% filter(!is.na(dep_delay),!is.na(arr_delay)) # Filter the flights not cancelled
  %>% group_by(year, month, day)
  %>% mutate(
    ranking = min_rank(desc(dep_time)),
    ranking_reverse = min_rank(dep_time),
    range_r = paste(range(ranking), collapse = " ~ ")
  )
  %>% filter(ranking %in% range(ranking))
  %>% arrange(year, month, day, ranking)
)
```

## Summary Mileage Each Day

```{r mileage-per-day}
count(flights, year, month, day, wt = distance)
```
```{r mileage-a-flight}
count(flights, year, month, day, tailnum, wt = distance)
```

## Popular Destinations

```{r popular-dests}
# Filter By Group Without Summary
flights %>%
  group_by(dest) %>%
  filter(n() > 365) %>%
  filter(arr_delay > 0) %>%
  mutate(prop_delay = arr_delay / sum(arr_delay)) %>%
  select(year: day, dest, arr_delay, prop_delay) %>%
  arrange(dest)
```

## Rolling up Summary

```{r rolling-up}
(daily <- group_by(flights, year, month, day))
(per_day <- summarise(daily, count = n()))
(per_month <- summarise(per_day, count = sum(count)))
(per_year <- summarise(per_month, count = sum(count)))
daily %>%
  ungroup() %>%
  summarise(flights = n())
```
## Distribution of Departure Times

```{r}
(flights_dt <- flights %>%
  filter(!is.na(dep_time), !is.na(arr_time)) %>% # Filter the flights not cancelled
  mutate(
    dep_time = make_datetime(year, month, day, dep_time %/% 100, dep_time %% 100),
    arr_time = make_datetime(year, month, day, arr_time %/% 100, arr_time %% 100),
    sched_dep_time = make_datetime(year, month, day, sched_dep_time %/% 100, sched_dep_time %% 100),
    sched_arr_time = make_datetime(year, month, day, sched_arr_time %/% 100, sched_arr_time %% 100)
  ) %>%
  select(origin, dest, ends_with("delay"), ends_with("time")))
flights_dt %>%
  filter(dep_time > ymd(20130104), dep_time < ymd(20130105)) %>%
  ggplot() + geom_freqpoly(aes(x = dep_time), binwidth = 60 * 10)
```

