---
title: "New York City Flights Data Playground"
output: html_notebook
---

```{r setup, include=FALSE}
library("dplyr")
```

## Dataset

```{r dataset}
flights
```
Take a simple summary of this dataset:
```{r summary_for_nycflights_dataset}
str(flights)
```

### Common Variable Types

Common Variable Types in DataFrame:
* `int` stands for integers
* `dbl` stands for doubles or real numbers
* `chr` stands for character vectors, where strings are included as well
* `dttm` is the abbreviation for `data-times`
* `date` stands for dates
* `lgl` stands for logical or vectors that contain `TRUE` or `FALSE`
* `fctr` stands for factors, which R uses to represent categorical variables with fixed possible values

## `dplyr` Functions

```{r filter_rows}
# Select all flights on Jan 1st.
filter(flights, month == 1, day == 1)
# Select all flights in Nov or Dec.
filter(flights, month == 11 | month == 12) # Pay attention to the `|` here, it is not `||`
filter(flights, month %in% c(9, 10))        # Select all flights in Sep or Oct.
filter(flights, between(dep_time, 0, 600))  # Departed between midnight and 6am
```
```{r near_usage_cases}
cat("[ sqrt(2) ^ 2 == 2 ] ->", sqrt(2) ^ 2 == 2)
cat("\n[ near(sqrt(2) ^ 2, 2) ] ->", near(sqrt(2) ^ 2, 2))
cat("\n[ 1/49 * 49 == 1 ] ->", 1/49 * 49 == 1)
cat("\n[ near(1/49 * 49, 1) ] ->", near(1/49 * 49, 1))
```
```{r De_Morgans_raw}
# Pay attention to the `|` here, which isn't `||`
all(
  filter(flights, !(arr_delay > 120 | dep_delay > 120)) ==
    filter(flights, arr_delay <= 120, dep_delay <= 120)
)
```
```{r missing_values}
NA
NA > 33
NA - 2
NA ^ NA
is.na(NA)    # A recommended way to determine whether a value is missing
```
`NA` will make almost any operation unknown.
```{r arrange_rows}
flights
arrange(flights, year, desc(month), day)
```
```{r select_columns}
select(flights, year:day)        # Same with the below
# select(flights, year, month, day)
select(flights, -(year: day))    # Select all colums except year, month and day
select(flights, starts_with("dep_"))
select(flights, contains("dep_"))
select(flights, contains("dep_"), everything()) # All dep fields moved to the start of the dataframe
select(flights, 出発時間 = dep_time, 予定出発 = sched_dep_time, 出発遅延 = dep_delay)
# Slight Difference between `select` and `rename`
rename(flights, 出発時間 = dep_time, 予定出発 = sched_dep_time, 出発遅延 = dep_delay)
```
```{r mutate_new_var}
mutate(
  select(flights, year:day, ends_with("delay"), distance, air_time),
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60,
  gain_per_hour = gain / air_time * 60 # Refer to the newly created column
)
transmute(
  flights,
  gain = dep_delay - arr_delay,
  speed = distance / air_time * 60,
  gain_per_hour = gain / air_time * 60 # Refer to the newly created column
)   # which will drop all original variables in the table
```
```{r Grouped_summaries}
summarise(flights, delay = mean(dep_delay, na.rm = TRUE), count = n())
summarise(
  group_by(flights, year, month, day),
  delay = mean(dep_delay, na.rm = TRUE),
  count = n()
)
```
```{r transformation_with_pipe}
# (delay_dataset <- summarise(
#   group_by(flights, dest),
#   count = n(),
#   dist = mean(distance, na.rm = TRUE),
#   delay = mean(arr_delay, na.rm = TRUE)
# ))
# (delay_filtered <- filter(delay_dataset, count > 20, dest != "HNL"))
#
# Use Pipe to make the transformation above concise:
(delays <- flights %>%
   group_by(dest) %>%
   summarise(
     count = n(),
     dist = mean(distance, na.rm = TRUE),
     delay = mean(arr_delay, na.rm = TRUE)
   ) %>%
   filter(count > 20, dest != "HNL")
)
ggplot(delays, aes(x = dist, y = delay)) +
  geom_point(aes(size = count), alpha = 1/3) +
  geom_smooth(se = FALSE)
```
```{r count_to_visualize_density_on_frequency_plot}
(delays <- flights
  %>% filter(!is.na(dep_delay),!is.na(arr_delay)) # Filter not cancelled flights
  %>% group_by(tailnum)
  %>% summarise(
    delay = mean(arr_delay),
    count = n()
  )
)
filter(delays, count > 25)
delays %>%
  ggplot(mapping = aes(x = delay)) +
  geom_freqpoly(binwidth = 10) +
  xlab("delay [min]")
delays %>%
  # filter(count > 25) %>%     # Filter to few flights to see variation
  ggplot(mapping = aes(x = delay, y = count)) +
  geom_point(alpha = 1/8) +    # Use the stastistic through count
  xlab("delay [min]")
```

