---
title: "Periodic Table Dataset Case"
author: "Richard Lea <chigix@zoho.com>"
date: "2018/12/12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r eval=FALSE}
download.file(url = "http://ismayc.github.io/periodic-table-data.csv",
 destfile = "../data-raw/periodic-table-data.csv")
```

Loading Data

```{r}
periodic_table <- read.csv("../data-raw/periodic-table-data.csv", stringsAsFactors = FALSE)
periodic_table
```

`View()` will open a new Pane to view this dataset:

```{r eval=FALSE}
View(periodic_table)
```

## Data Structures

### Data Frames

```{r}
str(periodic_table)
```

### Vectors

Create a **strings** vector and a **integer** vector:
```{r}
friend_names <- c("Ayanami", "Ikari", "Asuka", "Makinami")
friend_ages <- c(25L, 37L, 22L, 30L)
class(friend_names)
class(friend_ages)
```

* About the indicator of `L` on the difference between `integer` and `numeric`

```{r}
class(c(25,37,22,30))
class(seq(from = 0, to = 100, by = 2))
class(seq(from = 0L, to = 100L, by = 2L))
seq(from = 0L, to = 100L)
```

Using [colon operator](https://cran.r-project.org/doc/manuals/r-release/R-lang.html#Operators) to create sequence quickly:

```{r}
98:112
5:-5
```

Create Data frame with vectors previously created:

```{r}
friends <- data.frame(names = friend_names,
                      ages = friend_ages,
                      stringsAsFactors = FALSE`)
friends
```
`class(friends)` : `r class(friends)`

```{r}
str(friends)
```

### Factor

```{r}
table(periodic_table$block)
table(factor(periodic_table$block, levels = c("s", 'p', "d", "f")))
```

### Vectorized Operations

```{r}
seq(from = 2, to = 8)
seq(from = 2, to = 8) + 5L
periodic_table$name[10:20]
periodic_table$name[c(10,12)]
periodic_table$name[10:20][-c(1,2)]
```

### Filtering

```{r}
friend_names %in% c("Ayanami")
friend_names[friend_names %in% c("Ayanami")]
```

```{r}
periodic_table[(periodic_table$name %in% c("Hydrogen", "Oxygen")), 
               c("atomic_weight", "state_at_stp")]
```

```{r}
mean(123L)
```


